---
title: "Introduction to sparklyr"
output: html_document
---

We will largely follow Chapters 2 and 3 of Mastering Spark with R, [https://therinspark.com/](https://therinspark.com/).

## Installation

If you are using your own computer rather than Noteable, you may need to install Java 8 before you install Spark. Do it here:
[https://www.java.com/en/download/](https://www.java.com/en/download/).

I have commented out the line with spark_install('3.3'). This command installs Spark, and it only needs to be run once.

```{r echo = TRUE, warning=FALSE, message=FALSE}
library(sparklyr)
library(dplyr)
library(ggplot2)

# Install Spark
# This only needs to be run once
# spark_install('3.3')
```

## Connecting

You will be using the machine you are working on as a local cluster. This is just so you can familiarise yourself with using sparklyr. In practice, we would connect to an external cluster.

mtcars is a dataset that is built into R that we will use as an example.

```{r}
sc = spark_connect(master = "local")
cars = copy_to(sc, mtcars, overwrite =)
```

## Data input/output

### Write to a csv file:

This will create a folder in your working directory called cars.csv. It contains a csv with the cars data in it.

```{r}
spark_write_csv(cars, 'cars.csv')
```

Note that running this more than once will result in an error because spark_write_csv will not overwrite a folder you have already created. You may need to delete the folder cars.csv in your working directory in order to re-run.

### Read from a csv file:

```{r}
spark_read_csv(sc, 'cars.csv')
```

## Data wrangling

Familiar commands from dplyr work as you would expect, but now they instead connect to Spark and are run in parallel across the cluster.

### Create a new column.

```{r}
cars = mutate(cars, transmission = ifelse(am == 0, "automatic", "manual"))
```

### Select columns.

```{r}
select(cars, am, transmission) 
```

### Calculate the mean of each column.

```{r}
summarise_all(cars, mean, na.rm = TRUE)
```


## Plots
Creating a plot isn’t usually highly computationally demanding. Therefore, sparklyr does not have a full-fledged equivalent of ggplot. It is typically best to perform all data manipulations in Spark, then bring the result back to R using the collect() command. Finally, we use the regular ggplot package to make the graph.


```{r}
car_group = cars %>%
	group_by(cyl) %>%
	summarise(mpg = sum(mpg, na.rm = TRUE)) %>%
	collect()

ggplot(aes(as.factor(cyl), mpg), data = car_group) + 
  geom_col(fill = "#999999") + 
  coord_flip()
```


## Models, in brief

We will go into more details about what these models are in coming lectures.

### OLS

```{r}
ols_model = ml_linear_regression(cars, mpg ~ hp + disp)
summary(ols_model)
```

### Logistic regression

```{r}
lr_model = ml_logistic_regression(cars, am ~ hp + disp)
summary(lr_model)
```

### Multilayer perceptron

```{r}
mlp_model = ml_multilayer_perceptron_classifier(
	cars,
	am ~ hp + disp, 
	layers = c(2, 8, 8, 2))
predictions = ml_predict(mlp_model, cars)

select(predictions, prediction, probability_0, probability_1)
```

## Disconnecting

```{r}
spark_disconnect(sc)
```

